---
title: "Multiple comparisons & ANOVA, task N2"
author: "Natalia Sheiko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE,  
                      message = FALSE, 
                      warning = FALSE, 
                      results = 'hide')
```

```{r echo = F}
library(dplyr)
library(rstatix)
library(flextable)
library(multcomp)
library(ggplot2)
```

## A first glance at the data

```{r echo = FALSE, results='hold'}
soccer_general <- read.csv("soccer.csv", sep=";")[, 2:6] %>% 
    mutate(Position = as.factor(Position), 
    Nationality = as.factor(Nationality), 
    Age = as.numeric(Age), 
    Height = as.numeric(Height)
) %>% 
filter(Nationality %in% c("Spanish", "Italian", "German", "English", "Argentinian")) 

set.seed(1) 

soccer_wrk <- soccer_general[sample(1:nrow(soccer_general), 150), ] %>% 
    mutate(Nationality = factor(Nationality))

soccer_wrk %>% summary
```

## ANOVA for the differences among average height of soccer players on different field positions.

ANOVA requires three primary assumptions:

1.  The responses for each factor level have a normal distribution.
2.  The distributions for each factor level have the same variance.
3.  The data are independent.

**Note that players on one position do not provide information about players on other position, so data are independent.**

### Data imbalance checking & normality testing

**H0:** the distribution of the data is not significantly different from normal distribution.

**H1:** the distribution of the data is significantly different from normal distribution.

```{r results='hold', echo = F}
soccer_wrk %>%
  group_by(Position) %>%
  summarise(shapiro_test(Height), 
            length(Height),
            var(Height)) %>%
  mutate('hypothesis accepted' = ifelse(p.value > 0.05, 'H0', 'H1')) %>% 
  flextable() %>%
  set_table_properties(layout = "autofit") %>% 
  theme_box()
```

``` {r echo = F}

ggplot() +
  geom_boxplot(data = soccer_wrk, 
               aes(x = Height, y = Position)) +
  theme_light() +
  ggtitle("Comparison of players' height grouped by positions")

barplot(prop.table(table(soccer_wrk$Position)),
        main = "Class Distribution for players' field position")
```

### Homoscedasticity checking

**H0:** the variances are equal for all positions (groups).

**H1:** the variances are not equal for one pair of positions (groups) or more.

We use Bartlett test to determine whether or not the variances between several groups are equal.

```{r results='hold', echo = F}
bartlett.test(Height ~ Position, soccer_wrk)
```

Since the p-value is less than 0.05, we can reject the null hypothesis.

**Conclusion:**

1.  Based on the plot it clearly evident that near 25% of the data is in every class. There is no big difference observed in the amount of data available.

2.  We also can assume the normality for the players' height in all field position groups.

3.  Basing on the table above and Bartlett test's result we don't have sufficient evidence to say that the four groups have different variances.

***... therefore we can proceed to perform the one-way ANOVA.***

### One-way ANOVA

**H0:** there is no difference among players' height on different field positions

**H1:** there is a difference between players' height for one pair of field positions or more.

```{r results='hold', echo = F}
oneway.test(Height ~ Position, data = soccer_wrk)
```

Since the p-value is less than 0.05, we can reject the null hypothesis. 

There is a difference between players' height for some pair/pairs of field positions. To detect this pairs we can perform a post hoc analysis with Hothorn-Bretz-Westfall procedure

```{r results='hold', echo = F}
m0 <- lm(Height ~ Position, data = soccer_wrk)

HBW.comp <- m0 %>%  glht(linfct = mcp(Position = "Tukey"))
HBW.comp %>% summary()
HBW.comp %>% confint()
par(mar = c(5, 10, 4, 2)+0.1)
HBW.comp %>% plot(xlab = "Height difference (cm)")
```
So we can conclude that there is significant difference between height of players on this field positions:

   - Goalkeeper - Defender 
   - Midfielder - Defender
   - Goalkeeper - Forward 
   - Midfielder - Goalkeeper

The hometask N1 leads to another conclusion: there was significant differences only between this positions:

   - Defender - Forward
   - Forward - Midfielder

## ANOVA for the differences among average height of soccer players of different countries.

In general, the process is the same as in the example above.

**Players from one country do not provide information about  other players' countries, so data are independent.**

### Data imbalance checking & normality testing

**H0:** the distribution of the data is not significantly different from normal distribution.

**H1:** the distribution of the data is significantly different from normal distribution.

```{r results='hold', echo = F}

soccer_wrk %>%
  group_by(Nationality) %>%
  summarise(shapiro_test(Height), 
            length(Height),
            var(Height)) %>%
  mutate('hypothesis accepted' = ifelse(p.value > 0.05, 'H0', 'H1')) %>% 
  flextable() %>%
  set_table_properties(layout = "autofit") %>% 
  theme_box()
```

```{r results='hold', echo = F}
ggplot() +
  geom_boxplot(data = soccer_wrk, 
               aes(x = Height, y = Nationality)) +
  theme_light() +
  ggtitle("Comparison of players' height grouped by countries")

barplot(prop.table(table(soccer_wrk$Nationality)),
        main = "Class Distribution for players' country")
```

### Homoscedasticity checking

**H0:** the variances are equal for all positions (groups).

**H1:** the variances are not equal for one pair of positions (groups) or more.

We use Bartlett test to determine whether or not the variances between several groups are equal.

```{r results='hold', echo = F}
bartlett.test(Height ~ Nationality, soccer_wrk)
```
Since the p-value is more than 0.05, we can't reject the null hypothesis.

**Conclusion:**

1.  Based on the plot it clearly evident that near 20% of the data is in every class. So there is no big difference observed in the amount of data available.

2.  We can't assume the normality for one group of players' (Nationality = Spanish). This won't be a problem because the sample size had sufficient number of observations (50). We can assume the normality for other groups.

3.  Basing on Bartlett test's result we can say that the five groups have similar variance.

***... therefore we can proceed to perform general ANOVA (F-test).***

### ANOVA

**H0:** there is no difference among average height of players from different countries 

**H1:** there is a difference among average height of players from different countries 

```{r results='hold', echo = F}
aov(Height ~ Nationality, data = soccer_wrk) %>% summary
```

Since the p-value is more than 0.05, we can't reject the null hypothesis. 

```{r results='hold', echo = F}
m0 <- lm(Height ~ Nationality, data = soccer_wrk)
HBW.comp <- m0 %>%  glht(linfct = mcp(Nationality = "Tukey"))
HBW.comp %>% summary()
HBW.comp %>% confint()
par(mar = c(5, 10, 4, 2)+0.1)
HBW.comp %>% plot(xlab = "Height difference (cm)")
```
**Conclusion:**
We can conclude that there is no significant difference between height of players from different countries.