---
title: "Multiple comparisons, task N1"
author: "Natalia Sheiko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  
                      message = FALSE, 
                      warning = FALSE, 
                      results = 'hide')
```

```{r echo = FALSE}
library(dplyr)
library(ggplot2)
library(rstatix)
library(tidyr)
library(gtools)
library(flextable)
library(ggpubr)
```

## A first glance at the data

```{r echo = FALSE, results='hold'}
soccer_general <- read.csv("soccer.csv", sep=";")[, 2:6] %>% 
    mutate(Position = as.factor(Position), 
    Nationality = as.factor(Nationality), 
    Age = as.numeric(Age), 
    Height = as.numeric(Height)
) %>% 
filter(Nationality %in% c("Spanish", "Italian", "German", "English", "Argentinian")) 

set.seed(1) 

soccer_wrk <- soccer_general[sample(1:nrow(soccer_general), 150), ] %>% 
    mutate(Nationality = factor(Nationality))

soccer_wrk %>% summary
```

```{r echo = FALSE}
population_height <- mean(soccer_general$Height)

ggplot() +
  geom_boxplot(data = soccer_wrk,
               aes(x = Position, y = Height)) +
  geom_hline(yintercept = population_height, color = 'blue') +
  theme_light() + 
  ggtitle("Comparison of players' height grouped by positions with the population's mean")
  
```

## Confidence intervals for pairwise testing

```{r}
n = 6 # number of comparisons
FWER = 0.05
CL = (1-FWER)^(1/n) # Confidence level
```


```{r echo = FALSE, results='hold'}
pvalue_list <- list()

grid <- as.data.frame(combinations(n = 4, r = 2, v = as.character(unique(soccer_wrk$Position))))

mean_data <- soccer_wrk %>%
    group_by(Position) %>% summarise(mean(Height))

mean_data <- setNames(as.list(mean_data$`mean(Height)`), mean_data$Position)

for (number in 1:length(grid$V1)) {
  i <- grid[number, 'V1']   
  j <- grid[number, 'V2']
  grid[number, 'means diff'] <- round(as.numeric(mean_data[i]) - as.numeric(mean_data[j]), 2)
  sample <- soccer_wrk %>% filter(Position == i | Position == j)
  test_0 <- t.test(Height ~ Position, data = sample) 
  test <- test_0 %>% with(conf.int)
  grid[number,'CI'] <- paste(round(test[1], 2), round(test[2], 2), sep = ' - ')
  grid[number,'mean covering'] <- ifelse(grid[number, 'means diff'] <= round(test[2], 2), 'covers', 'does not cover')
  grid[number,'zero covering'] <- ifelse(round(test[2], 2) > 0 & round(test[1], 2) < 0, 'covers', 'does not cover')
  testB <- t.test(Height ~ Position, data = sample, conf.level = CL) %>% with(conf.int)
  grid[number,'simultaneous CI'] <- paste(round(testB[1], 2), round(testB[2], 2), sep = ' - ') 
  grid[number,'mean covering with Bonferroni correction '] <- ifelse(grid[number, 'means diff'] <= round(testB[2], 2), 'covers', 'does not cover')
  grid[number,'zero covering with Bonferroni correction'] <- ifelse(round(testB[2], 2) > 0 & round(testB[1], 2) < 0, 'covers', 'does not cover')
  pvalue_list <- append(pvalue_list,test_0$p.value)
}

grid %>% unite('pair', V1, V2, sep = '-') %>% flextable() %>%
  set_table_properties(layout = "autofit") %>% theme_box()

```

## Pairwise comparison

### Normality test

Paired t-test can be used only when the data is normally distributed. This can be checked using Shapiro-Wilk test.

H0: the distribution of the data is not significantly different from normal distribution. 

H1: the distribution of the data is significantly different from normal distribution.

```{r results = 'hold', echo = FALSE}
summary_table <- soccer_wrk %>% 
  group_by(Position) %>% 
  summarise(shapiro_test(Height)) %>%
  mutate('hypothesis accepted' = ifelse(p.value > 0.05, 'H0', 'H1')) %>% 
  flextable() %>%
  set_table_properties(layout = "autofit") %>% 
  theme_box()
summary_table
```


We can assume the normality for all the values in all groups, so pairwise t-test can be used.

### Pairwise test without correction:

```{r results = 'hold', echo = FALSE}
pw_test <- pairwise.t.test(soccer_wrk$Height, soccer_wrk$Position, 
                           pool.sd = FALSE, p.adjust.method = 'none')
pw_test

pw_test <- pairwise_t_test(soccer_wrk, Height ~ Position, 
                           pool.sd = FALSE, p.adjust.method = 'none') 

plot <- ggboxplot(soccer_wrk, x = "Position", y = "Height")
pw_test <- pw_test %>% add_xy_position(x = "Position")
plot + stat_pvalue_manual(pw_test, label = "p", label.size = 2.5)

```

### Pairwise test with correction:

#### Correction method: Holm

```{r results = 'hold', echo = FALSE}
pw_test_holm <- pairwise.t.test(soccer_wrk$Height, soccer_wrk$Position, 
                           pool.sd = FALSE)
pw_test_holm

pw_test_holm <- pairwise_t_test(soccer_wrk, Height ~ Position, 
                                pool.sd = FALSE) 

pw_test_holm <- pw_test_holm %>% add_xy_position(x = "Position")
plot + stat_pvalue_manual(pw_test_holm, label = "p.adj", label.size = 2.5)
```

#### Correction method: Benjamini & Hochberg

```{r  results = 'hold', echo = FALSE}
pw_test_hoch <- pairwise.t.test(soccer_wrk$Height, soccer_wrk$Position, 
                           pool.sd = FALSE, p.adjust.method = 'BH')
pw_test_hoch

pw_test_hoch <- pairwise_t_test(soccer_wrk, Height ~ Position, 
                                pool.sd = FALSE, p.adjust.method = 'BH') 

pw_test_hoch <- pw_test_hoch %>% add_xy_position(x = "Position")
plot + stat_pvalue_manual(pw_test_hoch, label = "p.adj", label.size = 2.5)
```

## Population vs sample

```{r results = 'hold', echo = FALSE}

grid_general <- as.data.frame(combinations(n = 4, r = 2, v = as.character(unique(soccer_general$Position))))

mean_data_general <- soccer_general %>%
    group_by(Position) %>% summarise(mean(Height))

mean_data_general <- setNames(as.list(mean_data_general$`mean(Height)`), mean_data_general$Position)

pw_test_holm_general <- pairwise_t_test(soccer_general, Height ~ Position, 
                                        pool.sd = FALSE) 
pw_test_hoch_general <- pairwise_t_test(soccer_general, Height ~ Position, 
                                        pool.sd = FALSE, p.adjust.method = 'BH') 

for (number in 1:length(grid_general$V1)) {
  
  i <- grid_general[number, 'V1']   
  j <- grid_general[number, 'V2']
   grid_general$`means difference in sample` <- grid$`means diff`
   grid_general[number, 'means difference in population'] <- round(as.numeric(mean_data_general[i]) - as.numeric(mean_data_general[j]), 2)
  sample <- soccer_general %>% filter(Position == i | Position == j)
  test_pvalue <- t.test(Height ~ Position, data = sample) %>% with(p.value)
  test_CI <- t.test(Height ~ Position, data = sample) %>% with(conf.int)
  grid_general[number,'CI for means difference in population'] <- paste(round(test_CI[1], 2), round(test_CI[2], 2), sep = ' - ')
  grid_general[number,'p-value for population'] <- format.pval(pv = test_pvalue, digits = 2, eps = 0.001, nsmall = 3)
 
}

grid_general %>%
  mutate('p-value, Holm correction' = format.pval(pv = pw_test_holm_general$p.adj, 
                                                  digits = 2, eps = 0.001, nsmall = 3)) %>%
  mutate('p-value, BH correction' = format.pval(pv = pw_test_hoch_general$p.adj,
                                                digits = 2, eps = 0.001, nsmall = 3)) %>%
  mutate("hypothesis accepted in sample's pairwise testing" = ifelse(pvalue_list > 0.05, 'H0', 'H1')) %>% 
  mutate("hypothesis accepted in population's pairwise testing" = ifelse(`p-value for population` > 0.05, 'H0', 'H1')) %>%
  flextable() %>%
  set_table_properties(layout = "autofit") %>% theme_box() %>% 
  style(i = 1,
        j = 9:10,
        pr_t = fp_text_default(italic = TRUE,
                               bold = TRUE,
                               color = "red"))

```

As we can see, there is a mismatching of hypotheses accepted between sample's pairwise testing and population's pairwise testing.

Therefore there is a significant difference between Defenders' and Forwards' height in population, that was missed in sample's pairwise testing.

The other hypotheses completely match for sample and population testing.
